---
title: "MATH250 Final Project - The Weeknd"
author: "Sergei Vorobev"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(here)
library(gt)

path <- here("Data", "the_weeknd.csv")
the_weeknd <- read_csv(path)
```

## Introduction

*The Weeknd* is a global superstar with incredible singing ability. His successful career boasts two Grammy wins, one for his performance of his song *Earned It* and another for his album *Beauty Behind the Madness*.[^1] Apart from the awards, he captures the hearts of many yearly, as he continues to be on Spotify's top 5 artists for a third year in a row now.[^2]

I was introduced to *The Weeknd* in middle school with his release of the album *Beauty Behind the Madness* and especially the song *The Hills*. It was popular at the time, and I really enjoyed its more pop take on the hip-hop genre. He quickly became one of my favorite artists and I followed all his subsequent releases. Since then my music taste has changed quite a lot, but *Kiss Land*, one of his early albums, remains my favorite.

*The Weeknd* sings about drug use and sex, which makes his music quite somber. In my opinion *Kiss Land* is incredible, surpassing the rest. Due to this, I want to see what makes this different and thus stand out. I also believe that `valence` and `danceability` are related in his music as I've seen a trend of his songs becoming more danceable, but less melancholy at the same time.

# Exploring tempo, danceability, and valence

```{r}
#| label: data cleaning

the_weeknd_sm <- the_weeknd %>% 
  select(album_name, track_name, tempo, danceability, valence) %>% 
  filter(album_name %in% c("Trilogy", 
                           "Kiss Land", 
                           "Beauty Behind The Madness",
                           "Starboy",
                           "My Dear Melancholy,",
                           "After Hours",
                           "Dawn FM"))
```

For this analysis I am only considering his studio album releases and one ep *My Dear Melancholy,* to avoid redundancy of deluxe and live versions, as well as re-releases.

```{r}
the_weeknd_albums <- the_weeknd_sm %>%
  group_by(album_name) %>% 
  summarise(Track_count = n(),
            Avg_tempo = round(mean(tempo), 3),
            Avg_danceability = round(mean(danceability), 3),
            Avg_valence = round(mean(valence), 3)) %>% 
  arrange(Avg_valence)
gt(the_weeknd_albums)
```

```{r}
#| label: first plot
#| fig-alt: A.

#ggplot()
```


[^1]: Bauer, Patricia. “The Weeknd | Biography, Songs, Albums, & Facts.” Encyclopedia Britannica, December 3, 2024. <https://www.britannica.com/biography/the-Weeknd>.

[^2]: Gavino, Nico. “Taylor Swift, the Weeknd and Bad Bunny Conquer Spotify'S 2024 Wrapped.” Hypebeast, December 4, 2024. <https://hypebeast.com/2024/12/taylor-swift-the-weeknd-and-bad-bunny-conquer-spotifys-2024-wrapped>.
